\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-11-10 icra-2021.tex
updated the intro

Planning a dynamic mission for unmanned aerial vehicles (UAVs) operating outdoors is a challenging task. Scenarios such as precision agriculture, search and rescue, and surveillance require advanced levels of autonomy but have limited energy budgets. A typical example of a mission with these limitations is a UAV flying a trajectory and performing some on-board computations. For instance, the UAV might detect some patterns and inform other, ground-based, actors. Planning a dynamic mission for such UAV involves finding optimal tradeoffs between the trajectory and computations. However, UAVs flying outdoors are often semi-autonomous, in the sense that the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems, and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy.

Planning algorithms for mobile robots are related to topics such as trajectory generation and path planning. Generally, these algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}, but in practice apply to a small number of robots~\cite{kim2005energy}, and focus on optimizing motion control for these robots~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being influenced by computations and not only bare motion~\cite{mei2005case}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}. Mission planning algorithms, which include a broader concept of a mission being a set of tasks along with a motion plan, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to a small number of robots~\cite{sadrpour2013mission,sadrpour2013experimental}. Given the availability of powerful GPU-equipped mobile hardware, the use of computations is expected to increase in the near future.

The user specifies an initial mission plan--the trajectory and the tasks--along with an energy budget and the following mission specification parameters. The Quality of Service (QoS) computations parameters are relative to the tasks, and trajectory-explicit equations (TEEs) adjustments parameters to the path. The proposed algorithm replans the initial mission plan dynamically as the energy budget changes due to atmospheric interferences. In the remainder of the paper, we adopt the following notation. We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, we refer to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and we refer to the current trajectory as TEE. Our goal is to complete the mission with the highest possible computations and adjustments as the UAV flies and its batteries drain. The algorithm both optimizes computations in the form of robot operating system (ROS) nodes, and it optimizes adjustments--a way to alter the trajectory--guiding the UAV using a vector field~\cite{de2017guidance} that converges smoothly to such trajectory. It relies on the assumptions of the mission being \emph{periodic} and \emph{uncertain}. Periodicity is often observed, e.g., when the UAV is flying in repetitive patterns for monitoring or search and rescue scenarios. Uncertainty accounts for the atmospheric interference, e.g., a fixed-wing UAV drifting due to windy weather. The algorithm addresses the periodicity by modeling the energy with Fourier analysis: the mission is periodic, so we expect the energy to also approximately evolve periodically. The uncertainty is addressed with a state estimator. The algorithm outputs the control input (adjustments and computations) using output model predictive control (MPC).

>>>>

Many scenarios involving unmanned aerial vehicles (UAVs) such as precision agriculture, search and rescue, and surveillance require high autonomy but have limited energy budgets. A typical example of a mission in these scenarios is a UAV flying a trajectory and performing some on-board computations with little human interaction. For instance, the UAV might detect ground patterns and inform other ground-based actors. Energy-aware dynamic mission planning finds optimal tradeoffs between the trajectory, computations, and energy requirements. Current generic planning solutions for UAVs flying outdoors do not plan the mission dynamically, nor are energy-aware. They are often semi-autonomous; the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy.

Planning algorithms literature for mobile robots is by now large and related to topics such as trajectory generation and path planning. Generally, the algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}. However, they apply to a small number of robots~\cite{kim2005energy}, and focus on planning the trajectory for these robots~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being influenced by computations~\cite{mei2005case}. Given the availability of powerful GPU-equipped mobile hardware, the use of computations is expected to increase in the near future. Mission planning algorithms, which include a broader concept of a mission being a set of tasks along with a motion plan, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to a small number of robots~\cite{sadrpour2013mission,sadrpour2013experimental}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}.

The proposed algorithm unfolds as follows. The user specifies an initial mission plan--the trajectory and the tasks--along with an energy budget and some parameters. The Quality of Service (QoS) computations parameters are relative to the tasks, and trajectory-explicit equations (TEEs) adjustments parameters to the path. The algorithm replans the initial mission plan as the energy budget changes due to atmospheric interferences. In the remainder of the paper, we adopt the following notation. We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, we refer to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and we refer to the current trajectory as TEE. Our goal is to complete the mission with the highest possible computations and adjustments as the UAV flies and its batteries drain.

The algorithm optimizes computations built as software components that run on a UAV's companion embedded device. Then, it optimizes adjustments--a way to alter the trajectory--and guides the UAV using a vector field~\cite{de2017guidance} that converges smoothly to such trajectory. It relies on the assumptions of the mission being \emph{periodic} and \emph{uncertain}. Uncertainty accounts for atmospheric interference. E.g., a fixed-wing UAV drifts due to windy weather. Periodicity accounts for the mission's repetitive patterns. Indeed UAV scenarios often iterate over a set of tasks and trajectories (e.g., a monitoring or search and rescue mission). The algorithm addresses the latter by modeling the energy with Fourier analysis. The mission is periodic, so we expect the energy to approximately evolve periodically. It addresses uncertainty with a state estimator and outputs the control input (adjustments and computations) using output model predictive control (MPC).
